#Couresera_rcode for Human Activity Recognition

##Summary
The data is about human activity recognition and is based on accelerometer and gyroscope data for 6 subjects.
A gradient boosted model is built.

##Input Data

###Reading Train Input

The training data set consits of 19623   records and the testing perediction data consists of 20 rows.
```{r}
traininput<-read.csv("E:\coursera\pml-training.csv") 
```


###Reading Test input
```{r}
testinput<-read.csv("E:\coursera\pml-testingna.csv") testinput$X<-NULL testinput$problem_id<-NULL
```

##Data Preprocessing

Unwanted information were removed .Missing value count for every column in testing data was found.Only columns in testing data that had had no missing values were considered.
Hence the total variables excluding the target was reduced from 169 to 58.
The same 58 variables were considerd in the training data along with the target.

###Understanding train data
```{r}
dim(traininput)
str(traininput)
str(traininput[,150:160]) 
classe_table<-table(traininput$classe)
 barplot(classe_table,main="Classe_count",xlab ="Classe",ylab="Count",col="blue")
 traininput[,1]<-NULL
 str(traininput)
```

###Taking count of null values in training data
```{r}
 nullcount<-colSums(is.na(traininput))
 barplot(nullcount,main="nullcount",xlab ="variablenames",ylab="nullcount",col="blue")
 plot(nullcount)
 variable_withnull<-which(nullcount>0)
 plot(variable_withnull) table(variable_withnull)
 barplot(variable_withnull,main="nullvar",xlab ="variablenames",ylab="nullcount",col="blue")
```

###Taking count of uniue values for every variable

```{r}
lu<-NULL 
for( i in 1:ncol(traininput)) 
{ 
lu[i]<-length(unique(traininput[,i]))
 } 
lu
names(traininput) 
variable_u<-cbind(names(traininput),lu)
str(variable_u) 
variable_u<-as.data.frame(variable_u)
```


###Taking null count of every column in testdata
```{r}
test_nullcount<-colSums(is.na(testinput)) 
test_withnull<-subset(test_nullcount,test_nullcount>0)
colnamesnotnull 
for(i in 1:ncol(testinput))
{ 
test_nullcount<-colSums(is.na(testinput)) 
colnamesnull<-names(test_withnull) 
}
```

###Considering only non null variables in test data(only around 58 variables are considred for model building)#### 
```{r}
test_nonull<-subset(test_nullcount,test_nullcount<1) 
testnonull<-names(test_nonull)
testp<-testinput[,testnonull]
```

###Taking same variables in trainingdata also
```{r}
traindata<-traininput[,testnonull] colnames(traindata)[colnames(traindata)=="traininput$classe"] <- "classe" str(traindata)
```


##Data Partioning.
The training data was partioned into sample training and sample testing data .The partition was based on the target variable such that 
same ratio of target classes were in both samples.
The training sample consists of   14718 rows   ,the testing sample consists of  4904 rows.

###Including required libraries
```{r}
library(caret) 
set.seed(1127)
```

###Split training data into training and testing data set
```{r}
inTraining <- createDataPartition(traindata$classe, p = .75, list = FALSE) trainingsample <- traindata[ inTraining,] testingsample <- traindata[-inTraining,]

fitControl <- trainControl(## 10-fold CV method = "repeatedcv", number = 10, ## repeated ten times repeats = 10)
```

##Model building 
A gbm model was built on the training sample and the accuracy was checked with the test sample and we obtained a accuracy of 97% by using confusion matrix.
```{r}
library(gbm)
set.seed(2029) 
gbmFit1 <- train(classe ~ ., data = trainingsample, method = "gbm", trControl = fitControl, ## This last option is actually one ## for gbm() that passes through verbose = FALSE)

gbmFit1
```

###Predicting the gbm fit model on sample test set obtained from the training partion and checking accuracy by confusion matrix####
Here, we get an accuracy of 97% on sample test data.

```{r}
testingsamplee<-testingsample[,1:58] 
tespred<-predict(gbmFit1,newdata=testingsamplee) 
confmatrix<-table( tespred,testingsample$classe)
confmatrix
```

###Predicting the gbm fit model on original testing data
```{r}
testset_prediction<-predict(gbmFit1,newdata=testp)
```

###Writing the result to a csv file
```{r}
write.csv(testset_prediction,"testset_prediction.csv")
```

The predicted output for test data set is as follows
B A B A A E D B A A B C B A E E A B B B